---
title: "SPM_2"
author: "Reagan Costello-White"
date: "6/16/2021"
output: html_document
---
# Load Packages

```{r, results = 'hide'}
library(tidyverse)
library(lubridate)
library(janitor)
library(readxl)
library(rmarkdown)
```

# Import raw data, turn into a tibble with snake_case names
# Mutate a new variable for fiscal year

```{r, results = 'hide'}
SPM2_FY15 <- read_excel("data/SPM2_FY15.xls", 
                        sheet = "Tab B - First Exit Detail") %>%
  as_tibble("SPM2_FY15.xls") %>%
  clean_names() %>%
  mutate(FY = 15)

SPM2_FY16 <- read_excel("data/SPM2_FY16.xls", 
                        sheet = "Tab B - First Exit Detail")%>%
  as_tibble("SPM2_FY16.xls") %>%
  clean_names() %>%
  mutate(FY = 16)

SPM2_FY17 <- read_excel("data/SPM2_FY17.xls", 
                        sheet = "Tab B - First Exit Detail")%>%
  as_tibble("SPM2_FY17.xls") %>%
  clean_names()%>%
  mutate(FY = 17)

SPM2_FY18 <- read_excel("data/SPM2_FY18.xls", 
                        sheet = "Tab B - First Exit Detail")%>%
  as_tibble("SPM2_FY18.xls") %>%
  clean_names()%>%
  mutate(FY = 18)

SPM2_FY19 <- read_excel("data/SPM2_FY19.xls", 
                        sheet = "Tab B - First Exit Detail")%>%
  as_tibble("SPM2_FY19.xls") %>%
  clean_names()%>%
  mutate(FY = 19)

SPM2_FY20 <- read_excel("data/SPM2_FY20.xls", 
                        sheet = "Tab B - First Exit Detail")%>%
  as_tibble("SPM2_FY20.xls") %>%
  clean_names()%>%
  mutate(FY = 20)

SPM2_FY21 <- read_excel("data/SPM2_FY21.xls", 
                        sheet = "Tab B - First Exit Detail")%>%
  as_tibble("SPM2_FY21.xls") %>%
  clean_names()%>%
  mutate(FY = 21)

demographics_FY13_FYTD21 <- read_csv("data/demographics_FY13_FYTD21.csv") %>%
    as_tibble("SPM2_FY21.xls") %>%
  clean_names()
```

# Merge SPM2 data frames then join demographics data

```{r}
d_all_FY <-  bind_rows(SPM2_FY15, SPM2_FY16, SPM2_FY17, SPM2_FY18, SPM2_FY19, SPM2_FY20, SPM2_FY21)
d_all_raw <- left_join(d_all_FY, demographics_FY13_FYTD21, by = c("client_uid" = "client_id"))
```

# Clean data file, change variable type, collapse factors and create variables
```{r, warning = FALSE}
d_all <-  d_all_raw %>% 
  group_by(client_uid) %>%
  arrange(client_uid) %>%
  mutate(race_f = factor(primary_race),
         proj_type_f = factor(proj_type),
         provider_f = factor(provider)) %>%
  mutate(race_fc = fct_collapse(race_f,
                                white = "White (HUD)",
                                black = "Black or African American (HUD)",
                                otherunknown = c("American Indian or Alaska Native (HUD)", 
                                                 "Native Hawaiian or Other Pacific Islander (HUD)",
                                                 "Asian (HUD)", "Client refused (HUD)", "Client doesn't know(HUD)",
                                                 "Data not collected (HUD)"))) %>%
  mutate(race_n = as.numeric(race_fc)) %>%
  mutate(end_date = as.Date(end_date, format = '%Y/%m/%d')) %>%
  mutate(days_cat = replace_na(days_to_reappear, 0)) %>%
  mutate(start_date = ymd(start_date),
         end_date = ymd(end_date)) %>%
           mutate(start_end_interval = interval(start_date, end_date),
                  start_end_days = time_length(start_end_interval, "day")) %>%
           mutate(year = year(end_date),
                  month = month(end_date)) %>%
           mutate(y_m = paste(year, month, sep = "-")) %>%
           relocate(client_uid, start_date, end_date, start_end_days, days_to_reappear,
                    race_fc, race_n) %>%
           slice(1) %>%
  ungroup
```

# Summary stats
#To create barplots with categorical variables, we first need to create a table of summary statistics

```{r}
sum_all <- d_all %>%
  select(client_uid, days_to_reappear, FY, race_fc, race_n, race_f, month, year)%>%
  na.omit()%>%
  summarise(n = n_distinct(client_uid), 
            mean = mean(days_to_reappear, na.rm = TRUE), 
            median = median(days_to_reappear, na.rm = TRUE),
            min = min(days_to_reappear, na.rm = TRUE), 
            max = max(days_to_reappear, na.rm = TRUE),
            sd = sd(days_to_reappear, na.rm = TRUE),
            sem = sd/sqrt(n()),
            upper_ci = mean + (1.96 * sem),
            lower_ci = mean - (1.96 * sem))%>%
  arrange(desc(n))

print(sum_all)
```

# Summary Table of Clients who returned to homelessness
```{r echo = FALSE, results='asis'}
library(gt)
gt_all <- gt(sum_all)
gt_all
```

# Table of clients by race
```{r}
sum_race <- d_all %>%
  select(client_uid, days_to_reappear, FY, race_fc, race_n, race_f, month, year)%>%
  na.omit()%>%
  group_by(race_f) %>%
  summarise(n = n_distinct(client_uid), 
            mean = mean(days_to_reappear, na.rm = TRUE), 
            median = median(days_to_reappear, na.rm = TRUE),
            min = min(days_to_reappear, na.rm = TRUE), 
            max = max(days_to_reappear, na.rm = TRUE),
            sd = sd(days_to_reappear, na.rm = TRUE),
            sem = sd/sqrt(n()),
            upper_ci = mean + (1.96 * sem),
            lower_ci = mean - (1.96 * sem))%>%
  arrange(desc(n))

print(sum_race)
```

```{r echo = FALSE, results='asis'}
library(gt)
gt_race <- gt(sum_race) %>%
  tab_header(title = "Mean and Median Days to Return to Homelessness by Race") %>%
  fmt_number(columns = 7:10, decimals = 2) %>%
  fmt_number(columns = 3, decimals = 2) %>%
  cols_width(upper_ci ~ px(100),
             lower_ci ~ px(100),
             n ~ px(50),
             mean ~ px(100),
             median ~ px(100),
             race_f ~ px(200)) %>%
  cols_align(align = "center") %>%
  cols_label(race_f = "Race",
             mean = "Mean",
             median = "Median",
             min = "Min",
             max = "Max",
             sd = "S. Dev",
             sem = "SEM",
             upper_ci = "Upper",
             lower_ci = "Lower") %>%
  tab_spanner(label = "95% Confidence Intervals",
              columns = c(upper_ci, lower_ci)) 

gt_race
```
# Table of clients by FY

```{r}
sum_FY <- d_all %>%
  select(client_uid, days_to_reappear, FY, race_fc, race_n, race_f, month, year)%>%
  na.omit()%>%
  group_by(FY) %>%
  summarise(n = n_distinct(client_uid), 
            mean = mean(days_to_reappear, na.rm = TRUE), 
            median = median(days_to_reappear, na.rm = TRUE),
            min = min(days_to_reappear, na.rm = TRUE), 
            max = max(days_to_reappear, na.rm = TRUE),
            sd = sd(days_to_reappear, na.rm = TRUE),
            sem = sd/sqrt(n()),
            upper_ci = mean + (1.96 * sem),
            lower_ci = mean - (1.96 * sem))%>%
  arrange(desc(FY))

print(sum_FY)
```

```{r echo = FALSE, results='asis'}
library(gt)
gt_FY <- gt(sum_FY) %>%
  tab_header(title = "Mean and Median Days to Return to Homelessness by Fiscal Year") %>%
  fmt_number(columns = 7:10, decimals = 2) %>%
  fmt_number(columns = 3, decimals = 2) %>%
  cols_width(upper_ci ~ px(100),
             lower_ci ~ px(100),
             n ~ px(50),
             mean ~ px(100),
             median ~ px(100),
             FY ~ px(100))%>%
  cols_align(align = "center") %>%
  cols_label(mean = "Mean",
             median = "Median",
             min = "Min",
             max = "Max",
             sd = "S. Dev",
             sem = "SEM",
             upper_ci = "Upper",
             lower_ci = "Lower",
             FY = "Fiscal Year") %>%
  tab_spanner(label = "95% Confidence Intervals",
              columns = c(upper_ci, lower_ci)) 

gt_FY
```